---
title: "Activity Montoring Data"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading and preprocessing the data
```{r loading, cache=TRUE}
activities <- read.csv("activity.csv", header=TRUE, sep=",", na.strings="NA")
activities$dt <- as.Date(activities$date)
summary(activities)
```

## What is mean total number of steps taken per day?
```{r activities}
library(ggplot2)
ggplot(activities, aes(dt, steps), fill=type) + geom_bar(stat = "identity") + labs(x="Date", y="Steps per Day", title = "Total Number of Steps Taken Per Day")
ggplot(activities, aes(x=dt, y=steps), fill=type) + stat_summary(fun.y="mean", geom="bar") + labs(x="date", y="Steps per Day (mean)", title = "Mean Number of Steps Per Day")
ggplot(activities, aes(x=dt, y=steps), fill=type) + stat_summary(fun.y="median", geom="bar") + labs(x="date", y="Steps per Day (median)", title = "Median Number of Steps Per Day")
myFun <- function(x) {
     c(mean = mean(x), median = median(x))
}
activityMeanMedian <- tapply(activities$steps, activities$dt, myFun)
head(activityMeanMedian)
```

## What is the average daily activity pattern?
```{r pattern}
library(plyr)
clean <- activities[!is.na(activities$steps),]
intervalTable <- ddply(clean, .(interval), summarize, Avg = mean(steps))
ggplot(intervalTable, aes(x=interval, y=Avg)) + geom_line() + labs(x="Interval", y="Average Steps", title="Average Number of steps")
#Maximum steps by interval
maxSteps <- max(intervalTable$Avg)
#Which interval contains the maximum average number of steps
intervalTable[intervalTable$Avg==maxSteps,1]
```

## Imputing missing values
#### 1. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NA
```{r missing}
sum(is.na(activities$steps))
```
#### 2. Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.
My Strategy will be filling NA with mean for that 5-minute interval
#### 3. Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r filling}
activitiesFiltered <- na.omit(activities)
meanByInterval <- aggregate(activitiesFiltered[, 1], list(activitiesFiltered$interval), mean)
colnames(meanByInterval) <- c("interval", "mean")
activitiesMerge <- merge(activities, meanByInterval, by="interval")
activitesNa <- is.na(activitiesMerge$steps)
activitiesMerge[activitesNa,"steps"] <- activitiesMerge[activitesNa, "mean"]
```

#### 4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day
```{r imputingHist}
ggplot(activitiesMerge, aes(dt, steps), fill=type) + geom_bar(stat = "identity") + labs(x="Date", y="Steps per Day", title = "Total Number of Steps Taken Per Day with Imputing Missing Values")
activityMeanMergeMedian <- tapply(activitiesMerge$steps, activities$dt, myFun)
head(activityMeanMergeMedian)
```
Do these values differ from the estimates from the first part of the assignment? 
Yes
What is the impact of imputing missing data on the estimates of the total daily number of steps?
mean increased and median decreased

## Are there differences in activity patterns between weekdays and weekends?
#### 1. Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
```{r weekdayweekend}
library(dplyr)
library(lubridate)
is_weekday <-function(date){
        if(wday(date)%in%c(1,7)) result<-"weekend"
        else
                result<-"weekday"
        result
}
activitiesFiltered <- mutate(activitiesFiltered,date=ymd(date)) %>% mutate(day=sapply(date,is_weekday))
table(activitiesFiltered$day)
```
#### 2. Make a panel plot containing a time series plot (i.e. type="l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). See the README file in the GitHub repository to see an example of what this plot should look like using simulated data.
```{r histweekendweekday}
intervalFilteredTable <- ddply(activitiesFiltered, .(interval, day), summarize, Avg = mean(steps))
ggplot(intervalFilteredTable, aes(x=interval, y=Avg)) + geom_line() + facet_grid(scales="free", space="free", .~day)
```